---
title: "MethylKit Parameter Testing"
author: "Yaamini Venkataraman"
date: "10/11/2018"
output: html_document
---

In this file, I'll try 1x, 3x, and 5x coverage with subset data from with [new `-score_min` parameter](https://yaaminiv.github.io/DML-Analysis-Part10/) to see how it affects clustering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
sessionInfo()
```

# Install packages

```{r}
install.packages("devtools") #Install the devtools package
library(devtools) #Load devtools

source("https://bioconductor.org/biocLite.R") #Source package from bioconductor
biocLite("methylKit") #Install methylkit
install_github("al2na/methylKit", build_vignettes = FALSE, repos=BiocInstaller::biocinstallRepos(), dependencies = TRUE) #Install more methylKit options
library(methylKit) #Load methylkit
```

# Process methylation data

I need to access the files on [gannett](http://gannet.fish.washington.edu/spartina/2018-10-10-project-virginica-oa-Large-Files/2018-10-03-Bismark-Parameter-Testing/2018-10-03-L012-Output/). The files could be downloaded using `curl`, but since gannet is mounted on the machine I'm using, I'll use an absolute path and access the files directly from gannet. This way, I don't have to write anything to my machine.

```{r}
analysisFiles <- list ("zr2096_1_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_2_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_3_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_4_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_5_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_6_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_7_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_8_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_9_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam",
                       "zr2096_10_s1_R1_bismark_bt2_pe.deduplicated.sorted.bam") #Put all .bam files into a list for analysis
```

